%%% paper.tex — PLOS 2026 submission
%%% ACM sigplan, anonymous, 10pt (per CFP)
%%% Max 6 pages (excl. bibliography), double-blind
\documentclass[sigplan,anonymous,10pt]{acmart}

% Disable CCS, keywords, and ACM ref format (not required for submission)
\settopmatter{printacmref=false}
\renewcommand\footnotetextcopyrightpermission[1]{}
\pagestyle{plain}

\begin{document}

\title{The Crux about Hardware Model Validations -- Towards Efficient Model Validation on Real Hardware}

\author{Author 1}
\affiliation{%
  \institution{Institution}
  \city{City}
  \country{Country}
}
\email{author1@example.com}

\author{Author 2}
\affiliation{%
  \institution{Institution}
  \city{City}
  \country{Country}
}
\email{author2@example.com}

%% ============================================================
%% Abstract
%% ============================================================
\begin{abstract}
System software is verified against formal models of hardware that essentially
provide a specification of the environment in which the software executes. The
validity of the verification result depends on the accuracy of the model.
%
Model validation---relating the specified behaviour with the observable
behaviour of real hardware---is hard, as modern processors execute many
invisible internal steps that influence the observable outcome, and worse, they
execute these steps speculatively.
%
We present a validation methodology that lets us validate real traces against a
formal model using a synthesis approach to reconstruct possible traces that
explain the observed behaviour on real hardware. To make this scale to
real-world examples, we present a reduction argument that constrains the
testing space while preserving representativeness.
%
We evaluate our approach on a model of the x86 memory-management unit~(MMU).
\end{abstract}

\maketitle

%% ============================================================
\section{Introduction}\label{sec:intro}
%% ============================================================

\subsection{Context}\label{sec:intro:context}

Verifying the correctness of application code has many facets. While the
application's own logic is an integral aspect, it is almost equally important
to ensure that the process environment is used correctly. This implies that we
need user-process-leverageable models of the process environment including both
the operating-system layer and the hardware layer.

The operating-system part of the process environment can be derived from a
reduction of a verified specification of the OS code, but the hardware model is
not as straightforward. Processor documentation can inform the design of a
hardware model; however, such documentation is often incomplete and existing
modelling practices may be incorrect. To ensure correct process-environment
specifications for applications to leverage, the underlying hardware model must
be tested and robust.

To this end, we propose an initial exploration of the different aspects
required to validate a hardware model and discuss how we apply these methods
to the x86 paging hardware.

\subsection{Problem}\label{sec:intro:problem}

Testing the hardware model comes with challenges unique to this scenario.
In traditional software verification, every action leading to a state
transition defined by the model is observable. In the case of a hardware
model, however, many events are triggered internally and are not directly
observable. We must therefore test that the model encompasses observed
behaviours using only externally visible actions, while anticipating the
internal events that can explain the behaviour seen.

For example, in the case of paging, a partial page walk being cached is not
directly observable. However, if this cached partial walk is later used to
complete a page walk and read data memory, and if the current page-table
writes do not match the observed read, then the caching can be inferred.

To assert that the model is correct, three aspects must be addressed:
\begin{enumerate}
  \item \textbf{Test-case generation.} We need exhaustive test cases that
        invoke all possible behaviours in the hardware.
  \item \textbf{Test execution harness.} We need to execute the tests with
        as little interference from other effects as possible.
  \item \textbf{Trace analysis.} We need to verify that the observed outcome
        is allowed by the proposed model.
\end{enumerate}
Addressing all three aspects ensures a robust hardware model.

%% ============================================================
\section{Motivation}\label{sec:motivation}
%% ============================================================

% TODO: Motivate why hardware model validation is critical.
% Discuss real-world consequences of incorrect models (e.g., soundness of
% verified system software depending on model fidelity).

%% ============================================================
\section{Background}\label{sec:background}
%% ============================================================

\subsection{Formal Hardware Models}\label{sec:bg:formal}

% TODO: Discuss existing formal hardware models and their role in
% software verification.

\subsection{Modern Hardware and Observable vs.\ Internal Events}%
\label{sec:bg:observable}

% TODO: Describe the distinction between externally observable events
% (e.g., memory reads/writes) and internal micro-architectural events
% (e.g., TLB fills, speculative walks, cache evictions).

\subsection{Software Synthesis}\label{sec:bg:synthesis}

% TODO: Introduce the synthesis problem and how it relates to
% reconstructing internal traces from observed behaviour.

%% ============================================================
\section{Proposed Model}\label{sec:model}
%% ============================================================

% TODO: Present the formal model of the x86 MMU that is the subject of
% validation. Describe states, transitions, and the distinction between
% external and internal actions.

%% ============================================================
\section{Reduction Argument}\label{sec:reduction}
%% ============================================================

The model is expected to be consistent across variables such as the number of
cores, virtual addresses, and physical addresses; however, testing across every
possible value is impractical. Large traces with many external and internal
actions cause state-space explosion and become computationally inviable.

To ensure efficiency without compromising correctness, we propose a set of
arguments that constrain the testing space while still representing any
possible configuration. This idea is inspired by the \emph{small-world
argument}: behaviours in the large state space can be represented by a
suitably chosen small subset of each parameter.

The aspects we consider are:
\begin{itemize}
  \item Number of cores
  \item Number of paging levels
  \item Number of virtual addresses
  \item Number of physical addresses for data
  \item Number of physical addresses for page-table entries
  \item Length of each trace
\end{itemize}

We present semi-formal arguments to constrain each of these variables.

As a proof of concept, we only consider legal page structures---we avoid
situations such as self-referencing PTEs and address aliasing---and we do not
consider huge pages.

\subsection{Cores}\label{sec:reduction:cores}

To determine the number of cores required, we first consider how the core
count affects the model. The state can be partitioned into
\emph{core-local state} (TLB, page-structure cache, store buffers, etc.)\ and
\emph{global state} (page-table and data memory). All cores share the global
state, and only the paging hardware updates it directly. Each core updates its
own core-local state.

Synchronisation effects arise from the interplay between a core updating its
local state and the hardware updating the global state (via page-table writes,
write-backs, evictions, etc.). Two cases must be tested:
\begin{enumerate}
  \item How do different changes in the global and local state affect
        behaviour within a single core?
  \item Do global updates triggered by local core updates follow total store
        order as expected by the model?
\end{enumerate}

For the first case, one additional core suffices to induce global changes,
giving a minimum of \textbf{two cores}. For the second case, detecting cycles
in read--write anti-dependencies also requires only two cores in principle,
though four cores (two writers and two readers) make the effect easier to
observe in practice.

\subsection{Paging Levels}\label{sec:reduction:levels}

Traditional x86 uses four-level paging
(PML4~$\to$ PDPT~$\to$ PD~$\to$ PT~$\to$ Page). We classify levels into
equivalence classes: \emph{root} (PML4), \emph{intermediate} (PDPT, PD), and
\emph{leaf} (PT). The root is unique because no other link points to it; the
leaf is unique because its entries carry distinct permission bits and point to
data pages.

We need enough levels to expose the non-atomic nature of the page walk.
One level makes the walk atomic and is insufficient. Two levels
(PML4~$\to$ PD~$\to$ PT) allow non-atomic behaviour for 4\,KB pages and
cover all three equivalence classes. Each link variation can be decomposed into
effects of individual links, so two levels encompass all interesting cases.

We achieve two-level paging for 4\,KB pages by fixing the higher levels.

\subsection{Physical PTE Addresses}\label{sec:reduction:pte}

Different physical addresses represent the same entity, so repeating a test
with different addresses does not present new behaviour. To determine the
required number, we consider the maximum needed to represent a test scenario
that cannot be decomposed into smaller tests.

Physical addresses can be classified into those within the same page and those
on different pages. Starting from a single page-aligned address, we need at
least one intra-page address and one inter-page address, yielding a small
finite set.

% TODO: Finalise the concrete parameter values.

%% ============================================================
\section{Test-Case Generation}\label{sec:testgen}
%% ============================================================

The test cases take the form of traces of externally invocable actions---e.g.,
writing a page-table entry, reading data memory---that correspond to the
externally visible actions in the model. Rather than hand-crafting litmus
tests, we automate the generation of a sufficiently exhaustive set of traces
encompassing all possible behaviours, using the reduced model from
\cref{sec:reduction}.

% TODO: Describe the generation algorithm and coverage criteria.

%% ============================================================
\section{Testing Architecture}\label{sec:harness}
%% ============================================================

To execute the test traces and obtain corresponding results, we built a test
harness using the Barrelfish operating system, ensuring execution as close to
the hardware as possible.

% TODO: Describe the harness architecture, how hardware state is
% prepared (e.g., PWC entries), and how interference is minimised.

%% ============================================================
\section{Trace Analysis}\label{sec:analysis}
%% ============================================================

We approach verification of a trace against the model as a \emph{synthesis
problem}. The model dictates the constraints, derived from states and
transitions, and the externally observed events (together with the outputs
from test execution) guide generation of potential internal-state transitions
that justify the observed behaviour. When no such completion exists, the model
is violated, directing its improvement.

% TODO: Formalise the synthesis encoding and discuss scalability.

%% ============================================================
\section{Evaluation}\label{sec:eval}
%% ============================================================

% TODO: Present evaluation results.
% - Number/type of generated test cases
% - Execution statistics (time, hardware platform)
% - Model violations found and model improvements made
% - Scalability of the synthesis step

%% ============================================================
\section{Related Work}\label{sec:related}
%% ============================================================

% TODO: Discuss related work on hardware model validation, memory-model
% testing (e.g., litmus tests), TLB/MMU modelling, and synthesis-based
% verification.

%% ============================================================
\section{Conclusion}\label{sec:conclusion}
%% ============================================================

% TODO: Summarise contributions and outline future work.

%% ============================================================
%% Bibliography
%% ============================================================
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
