%% ============================================================
\section{Introduction}\label{sec:intro}
%% ============================================================

\subsection{Context}\label{sec:intro:context}

To be able to verify the correctness of an application’s code has 
many aspects. While the application’s own code is an integral aspect, 
it is almost equally important for the application code to ensure that 
the process environment is used correctly. This implies that we need 
user process leverageable models of the process environment including 
the operating system part and the hardware part. While the operating 
system model part of the process environment can be derived from a 
reduction of a verified specification of the operating system code, the 
hardware model is not as straightforward. Documentation can help in 
proposing a hardware model however, these are often lacking and 
existing practices may be incorrect. Thus, to ensure correct process 
environment specifications for applications to leverage, it is 
essential that it is based upon a tested and robust hardware model. To 
this end, we propose an initial exploration of the different aspects 
required to validate a hardware model. We discuss how we use these 
methods and our observations for the x86 paging hardware.

\subsection{Problem}\label{sec:intro:problem}

Testing the hardware model comes with challenges that are unique to 
this scenario: In verifying software traditionally, we are able to 
observe and identify every action that leads to a transition in state 
defined by our model. In the case of the hardware model however, there 
are many events that are triggered and happen within the hardware that 
are not directly observable. Thus, we need to test that our model 
encompasses observed behaviours using externally visible actions and 
anticipating the internal events that can explain the behaviour seen. 
For eg: in the case of paging, a partial page walk being cached is not 
observable, however if this cached page walk is used later to complete 
the page walk and read data memory, then depending on if the current 
page writes do not match the observed read, the caching can be inferred.

To be able to assert that our model is correct, there are 3 aspects to 
consider: 1) we need exhaustive test cases that invoke all possible 
behaviours in the hardware, 2) Using these test cases, we need to find 
a way to test the hardware with as little interference from other 
effects as possible, 3) We also need to verify that the observed 
outcome from the tests is allowed by our proposed model.  Addressing 
all of these would ensure that we have a robust hardware model.

To assert that the model is correct, three aspects must be addressed:
\begin{enumerate}
	\item \textbf{Test-case generation.} Generate interesting test 
cases (how can we get stimulus that can expose the interesting behavior 
to validate the model)

	\item \textbf{Test execution harness.} Test execution harness 
(how can we execute the test such that the desired behavior is actually 
exposable) with how do we get the hardware into a state for the tests 
(e.g., how to set the PWC entries accordingly)

	\item \textbf{Trace analysis.} Checking efficiently for 
explainability. (How do we validate the observed trace against the 
model (i.e., finding the hidden steps efficiently))

\end{enumerate}
Addressing all three aspects ensures a robust hardware model.

\subsection{Approach}\label{sec:intro:approach}


\subsection{Results}\label{sec:intro:results}

% TODO: What are our results?
