%% ============================================================
\section{Motivation}\label{sec:motivation}
%% ============================================================

% TODO: Motivate why hardware model validation is critical.
% Discuss real-world consequences of incorrect models (e.g., soundness of
% verified system software depending on model fidelity).
%   Thus specifications that expose the process environment usage are essential. While these specifications are partly derived from the operating system kernel implementation, another major part is the hardware itself. The process environment specification integrally depends on a model of the hardware and if this is incorrect, it results in incorrect and ineffective specifications.To this end, we persevere to propose a robust consideration of the different aspects of hardware model creation and testing. 
% One might assume that we can use existing documentation to create a model however, often these documentation are incomplete and do not cover various corner cases and interaction of effects. Similarly, using existing operating systems as a baseline to understand how they interact with the hardware is also not a robust enough method since these assumptions may not be completely accurate and lead to errors. 
Thus, although there are sources that can guide us in creating an anticipated model of the hardware, it is essential for it to be tested against a reasonable set of inputs ensuring that it is actually correct. The focus of our paper is to highlight the various aspects of this process to be considered and ways we have experimented in controlling this process to be viable.

Proofs are only as sound as the environment they assume. Application proofs assume OS and hardware behaviours which are encoded as formal models. Therefore, verification is only as sound as the environment model: unsound hardware models can invalidate higher-level proofs. This section makes that dependency explicit, clarifies why hardware is an unusually hard object to model, and motivates validating the model itself rather than only using it as a trusted oracle.

\paragraph{Verification chain of trust.}
Many existing works that deal with verification in systems focus primarily on verifying the Operating System Kernel itself or the Application code. But these two pieces are not sufficient since correctness of OS code and Application code does not ensure interactions of the application with the process environment is correct. They blindly depend on specifications of the process environment: system calls, memory management, and the instruction set architecture as realized by the processor. Those specifications in turn assume a model of hardware behaviour, often including microarchitectural effects that matter for paging, caching, and coherence. The chain is therefore hierarchical: application proofs assume OS abstractions, OS proofs assume hardware, and hardware behaviour is ultimately captured only through documentation, prior experiments, existing practices and engineered formal models. An unsound or incomplete hardware model breaks this chain: a formally verified operating system built atop an inaccurate MMU model may still exhibit behaviours not captured by the proof, because the proof's world and the machine's world diverge. Our concern is precisely the fidelity of that lowest formal link when modelling paging-related hardware.

\paragraph{Difference between software and hardware verification.}
Software verification typically allows for explicit transitions, inspectable state, and reconstructible control flow, even when abstraction is heavy. Hardware validation lacks this advantage: many transitions are invisible and are often inferable only indirectly from what is exposed at the boundary. Hardware validation is fundamentally a partial observability problem. Processors also exacerbate this problem due to speculative actions that may not be observable. There are also many ambiguous traces that can be explained by multiple internal executions. Thus, this leads us to a synthesis-style reasoning over hidden steps to identify whether there
exists a sequence of admissible steps consistent with both the
model and the measurements which is discussed in more detail in the section \ref{sec:bg:synthesis}. 

\paragraph{Why documentation existing practices are inadequate.}
Using existing documentation to create a model is ineffective since often these documentation are incomplete and do not cover various corner cases and interaction of effects. Relying on
existing operating systems as a proxy is similarly fragile: kernels
encode assumptions or obsolete expectations, depend on empirically discovered behaviour, and ship workarounds for undocumented effects. None of that constitutes a guarantee of hardware semantics. Thus more robust validation is essential.
